TensorAlloy: an automatic atomistic neural network program for alloys

### 2019-08-06 doc/papers/nn/manuscript.tex
Final revision for this submission.

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. Currently, the symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help.  
In this work, we propose a virtual-atom approach to model the total energy of 
symmetry function descriptors based atomistic neural network. Our approach 
starts from atomic positions directly so that the derivations of forces and 
virial can be handled by TensorFlow automatically and efficiently. This new 
approach is implemented in our program TensorAlloy, which supports constructing 
machine learning interaction potentials for both \textit{molecules} and 
\textit{solids}. The QM7 and SNAP/Ni-Mo datasets are used to demostrate the 
performances of our program.
\end{abstract}

@@

In order to precisely model atomistic interactions, both total energy and atomic 
forces are necessary. For \textit{solids}, the virial stress tensor is also an
essential metric. However, integrating force and stress into the loss function 
is not an easy task. The machine learning approaches favor vectorizable 
operations but the calculations of atomic descriptors are typically complicated 
so that many traditional ML packages (Amp, kCON, etc) choose to pre-compute 
descriptors and only build computation graph from descriptors to total energy. 
This traditional approach can significantly reduce technical difficulty and 
works very well if only total energy is considered in the loss function. But to 
fit atomic forces, numerous codes must be written manually, such as the 
derivation of descriptors with respect to positions. Although the derivation is 
theoreticallyclear, the implementation is challenging and the efficiency can not 
be guaranteed.

@@

In the atomistic neural network (\textbf{ANN}) framework, the total energy, 
$E^{total}$, of a structure with $N$ atoms, is the sum of all atomic energies:
\begin{equation}
\label{eq:general_e_total}
E^{total} = \sum_{i}^{N}{E_i}
\end{equation}
where $E_{i}$, the energy of atom \textit{i}, is the output of the neural 
network $\mathbf{NN}_{el}$ for element $el$:
\begin{equation}
E_i = \mathbf{NN}_{el}\left( \mathbf{G}_i \right)
\end{equation} 
Here $\mathbf{G}_i$ is a vector representing atomic descriptors of atom $i$. 
$\mathbf{G}_i$ typically only depends on local environments, i.e.  
neighbors of atom $i$:
\begin{equation}
\mathbf{G}_i = F_G(\left\{ \rijn, |\rijn| < r_c\right\} )
\end{equation}
where $F_G$ can be arbitrary function, $r_c$ is the cutoff radius and $\rijn$ is 
the interatomic distance considering the periodic boundary conditions: 
\begin{equation}
    \label{eq:rijn}
    \rijn = \Vert \mathbf{r}_i^{(0)} - \mathbf{r}_j^{(0)} + 
    \mathbf{n}^T \mathbf{h} \Vert
\end{equation}
where $\mathbf{r}_i^{(0)}$ and $\mathbf{r}_j^{(0)}$ denote positions of $i$ and 
$j$ in the primitive cell, $\mathbf{n}$ is a column vector specifying the cell 
shifts along $(X,Y,Z)$ directions 
\begin{equation}
    \label{eq:ij_shift}
    \mathbf{n} = \begin{pmatrix}
        n_x \\
        n_y \\
        n_z
    \end{pmatrix}
\end{equation}
and $\mathbf{h}$ is the \textbf{row-major} $3 \times 3$ lattice tensor:
\begin{equation}
    \label{eq:lattice}
    \mathbf{h} = \begin{pmatrix}
        h_{xx} & h_{xy} & h_{xz} \\
        h_{yx} & h_{yy} & h_{yz} \\
        h_{zx} & h_{zy} & h_{zz} \\
    \end{pmatrix}
\end{equation}

@@

Then we can determine another two key constants: $\nijmax$ and $\nijkmax$. 
Here $\nijmax$ represents the maximum number of atom-atom pairs in
arbitrary structure and the subscript 'ijk' denotes triples 
(or three-atoms interactions). $\nijmax$ and $\nijkmax$ only depend on 
$r_{cut}$ and the associated codes are given in Appendix(B). 
Table \ref{table:nij_nijk_max} summarizes $\nijmax$ and $\nijkmax$ at different
$r_{cut}$ of the two demo datasets: QM7 and Ni-Mo (SNAP).

>>>>

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. The symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help.  
In this work, we propose a virtual-atom approach to model the total energy of 
symmetry function descriptors based atomistic neural network. Our approach 
creates the computation graph directly from atomic positions. Thus, the 
derivations of forces and virial can be handled by TensorFlow automatically and 
efficiently. This new approach is implemented in our program TensorAlloy, which 
supports constructing machine learning interaction potentials for both 
\textit{molecules} and \textit{solids}. The QM7 and SNAP/Ni-Mo datasets are used 
to demostrate the performances of our program.
\end{abstract}

@@

In order to precisely model atomistic interactions, both total energy and atomic 
forces are necessary. For \textit{solids}, the virial stress tensor is also an
essential metric. However, integrating force and stress into the loss function 
is not an easy task. The machine learning approaches favor vectorizable 
operations but the calculations of atomic descriptors are typically complicated 
so that many traditional ML packages (Amp, kCON, etc) choose to pre-compute 
descriptors and only build computation graph from descriptors to total energy. 
This traditional approach can significantly reduce technical difficulty and 
works very well if only total energy is considered in the loss function. But to 
fit atomic forces, numerous codes must be written manually, such as the 
derivation of descriptors with respect to positions. Although the derivation is 
theoretically clear, the implementation is challenging and the efficiency can 
not be guaranteed.

@@

In the atomistic neural network (\textbf{ANN}) framework, the total energy, 
$E^{total}$, of a structure with $N$ atoms, can be expressed as the sum of all 
atomic energies:
\begin{equation}
\label{eq:general_e_total}
E^{total} = \sum_{i}^{N}{E_i}
\end{equation}
where $E_{i}$, the energy of atom \textit{i}, is the output of the neural 
network $\mathbf{NN}_{el}$ for element $el$:
\begin{equation}
E_i = \mathbf{NN}_{el}\left( \mathbf{G}_i \right)
\end{equation} 
Here $\mathbf{G}_i$ is a vector representing atomic descriptors of atom $i$. 
$\mathbf{G}_i$ typically only depends on local environment of atom $i$, i.e.  
its neighbors:
\begin{equation}
\mathbf{G}_i = F_G(\left\{ \rijn, |\rijn| < r_c\right\} )
\end{equation}
where $F_G$ is a specifc feature function, $r_c$ is the cutoff radius and 
$\rijn$ is the interatomic distance considering the periodic boundary 
conditions: 
\begin{equation}
    \label{eq:rijn}
    \rijn = \Vert \mathbf{r}_i^{(0)} - \mathbf{r}_j^{(0)} + 
    \mathbf{n}^T \mathbf{h} \Vert
\end{equation}
where $\mathbf{r}_i^{(0)}$ and $\mathbf{r}_j^{(0)}$ denote positions of atoms 
$i$ and $j$ in the primitive cell, $\mathbf{n}$ is a column vector specifying 
the cell shifts along $(X,Y,Z)$ directions
\begin{equation}
    \label{eq:ij_shift}
    \mathbf{n} = \begin{pmatrix}
        n_x \\
        n_y \\
        n_z
    \end{pmatrix}
\end{equation}
and $\mathbf{h}$ is the \textbf{row-major} $3 \times 3$ lattice tensor:
\begin{equation}
    \label{eq:lattice}
    \mathbf{h} = \begin{pmatrix}
        h_{xx} & h_{xy} & h_{xz} \\
        h_{yx} & h_{yy} & h_{yz} \\
        h_{zx} & h_{zy} & h_{zz} \\
    \end{pmatrix}
\end{equation}

@@

Then we can determine another two key constants: $\nijmax$ and $\nijkmax$. 
Here $\nijmax$ represents the maximum number of atom-atom pairs in
arbitrary structure and the subscript 'ijk' denotes triples 
(or three-atoms) interactions. $\nijmax$ and $\nijkmax$ only depend on 
$r_{cut}$ and the associated codes are given in Appendix(B). 
Table \ref{table:nij_nijk_max} summarizes $\nijmax$ and $\nijkmax$ at different
$r_{cut}$ of the two demo datasets: QM7 and Ni-Mo (SNAP).
